---
title: "Assignment"
subtitle: "Option Pricing and Risk - Financial Statistics"
author: "Alexander J Ohrt"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    theme: readable
    highlight: textmate
    number_sections: true
  pdf_document:
    fig_caption: true
    number_sections: true
    toc: false
editor_options: 
  chunk_output_type: console
geometry:
  margin = 2.2cm
urlcolor: blue
linkcolor: black
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = T, warning = F, fig.width = 10, comment = "#>", size = "scriptsize")
setwd("/home/ajo/gitRepos/finance/part1")
library(ggplot2)
library(tidyverse)
```

# Problem A - Euler-Maruyama Method

First we implement a discretization for the Black-Scholes-Merton price dynamics with the given parameter set, with $n = 250$ steps. We generate $M_1 = 10$, $M_2 = 100$, $M_3 = 1000$, $M_4 = 10000$ and $M_5 = 100000$ paths and plot the paths in separate figures. 

```{r}
set.seed(061999)
s0 <- 12
T1 <- 1
mu <- 0.03
sigma <- 0.17
n <-  250
t <- seq(0, T1, length.out = n)
M1 <- 10
M2 <- 100
M3 <- 1000
M4 <- 10000
M5 <- 100000
```

```{r}
# Price path stochastic process. 
price.path <- function(s0, t, mu, sigma){
  W <- rnorm(n = length(t), mean = 0, sd = t)
  s0*exp(mu*t)*exp(sigma*W-sigma^2/2*t)
}
```

```{r}
#plot(NULL, NULL, main = paste0(M1," paths"), xlim = c(0,1), ylim = c(0, 50), ylab = "S", xlab = "t")
M1.paths <- matrix(rep(NA, length.out = M1*n), nrow = M1)
for (i in 1:M1){
  M1.paths[i,] <- price.path(s0, t, mu, sigma)
}
#lines(M1.paths, col = 1:M1)
#df1 <- tibble(M1.paths)
#df1 %>% ggplot()

# Burde lage denne plotting bedre med ggplot senere kanskje!
# Endre til tid på x-aksen også kanskje!
matplot(t(tibble(M1.paths)), type = "l", main = paste0(M1," paths"), xlab = "Discretization Points", ylab = "S")
```

```{r}
#plot(NULL, NULL, main = paste0(M1," paths"), xlim = c(0,1), ylim = c(0, 50), ylab = "S", xlab = "t")
M2.paths <- matrix(rep(NA, length.out = M2*n), nrow = M2)
for (i in 1:M2){
  M2.paths[i,] <- price.path(s0, t, mu, sigma)
}
#lines(M1.paths, col = 1:M1)
#df2 <- tibble(M2.paths)
#df1 %>% ggplot()

# Burde lage denne plotting bedre med ggplot senere kanskje!
# Endre til tid på x-aksen også kanskje!
#M2.p <- tibble(M2.paths)
matplot(M2.paths, type = "l", main = paste0(M2," paths"), xlab = "Discretization Points", ylab = "S")
lines(t, apply(M2.paths, 2, mean), cex = 2)
```

Dette skal gjøres for alle verdiene for M over! (gjør etter at plottingen er fikset!)

TESTING CODE BELOW COMPARING TO SLIDE 33. Tror noe er galt!

```{r}
n2 <- 100
Mtest.paths <- matrix(rep(NA, length.out = M1*n2), nrow = M1)
for (i in 1:M1){
  Mtest.paths[i,] <- price.path(10, seq(0, 5, length.out = n2), 
                                0.05, 0.05)
}
#lines(M1.paths, col = 1:M1)
#df2 <- tibble(M2.paths)
#df1 %>% ggplot()

# Burde lage denne plotting bedre med ggplot senere kanskje!
# Endre til tid på x-aksen også kanskje!
matplot(t(tibble(Mtest.paths)), type = "l", main = paste0(M1," paths"), xlab = "Discretization Points", ylab = "S", ylim = c(0,20))
```

ÅPENBART AT NOE ER GALT HER!

Prøvde å få denne til å ligne på plot på slide 33 i Section 2.2, men mine paths varierer veldig mye! Tror det er noe galt i koden for pathsene muligens!!

The Monte Carlo estimator for $\hat{S}_T$ is calculated separately for each of the values of $M_i$. The 5\% confidence interval is provided for each estimator. A comparison between these estimators and the analytical solution of $\mathbb{E}(S_T)$ (seen above) is done and differences are explained. 

# Problem B - Option Pricing I

```{r}
set.seed(061999)
s0 <- 24
T1 <- 1.5
r <- 0.02
sigma <- 0.22
K <- 23.5
```


```{r, eval = F}
# BSM pricing formula for European call option, 
# to calculate the fair price at t = 0 in closed form.
BSM <- function(s0, T1, r, sigma, K){
  d1 <- (log(s0/K)+(r+sigma^2/2)*T1)/(sigma*sqrt(T1))
  print(d1)
  d2 <- d1 - sigma*sqrt(T1)
  print(d2)
  print(pnorm(d1))
  print(pnorm(d2))
  s0*pnorm(d1) - exp(-r*T1)*K*pnorm(d2)  
}

# Test with result on slide 48.
BSM(35, 0.5, 0.05, 0.25, 35.5)
# Looks correct!^
```

We implement a Monte Carlo estimator for this option price by simulation paths with the Euler-Maruyama method for given steps and paths. 

# Problem C - Option Pricing II

A Monte Carlo estimator is implemented for a specific Asian Call Option. 

```{r}
set.seed(061999)
```
